\chapter*{Acknowledgements}

There are many people without whom the experience of creating this thesis would have been made more challenging, if not impossible. In lieu of a perfect and complete list, which no doubt could be a full text of its own, I offer this imperfect and incomplete substitution.

At a fundamental level, I would like to thank the people who made this work possible in the first place: The Quantum Engineering Centre for Doctoral Training (QECDT) for offering me a studentship, the Quantum Engineering Technology Labs (QETLabs) for offering me a desk, and the Heilbronn Institute for Mathematical Research for funding my PhD. I also thank my supervisors -- Peter Turner and Noah Linden -- for keeping my work on track, my other collaborators -- Ashley Montanaro, Ra\'{u}l Garc\'{i}a-Patr\'{o}n and Jelmer Renema -- for such great ideas, discussions and opportunities, and Ryan Mann for helpful comments on this thesis.

I have worked alongside many great fellow researchers during my time in the above groups. Of particular note are my fellow QECDT students: Daniel Love, Geraint Gough, Henry Semenenko, Jason Mueller, Joe Smith, Lawrence Rosenfeld, Lucio Stefan, Martin Nicolle and Sam Holder. I also thank other Bristol theorists, including Lana Mineh Jake Bulmer, Oli Thomas, Rachel Chadwick, Stasja Stanisic, Sam Morley-Short, Stephen Piddock and Sam Pallister. It was a pleasure working with all of you and I wish you the best of luck in your future endeavours.

Thank you to the many people I have worked with on outreach, both in QECDT and QETLabs but also outside, such as in Pint of Science and with Bristol Doctoral College. I would in particular like to thank Nic Harrigan, Euan Allen, Alasdair Price, Holly Caskie, Jamie Thakrar, Sophie Stephens, Laura Veldenz, Ben Barber and Kate Oliver, for the opportunities they have given me to act a fool in the name of public engagement.

Of course, University is only one part of the rich tapestry that is a PhD life. Outside of University, I would like to thank the many sporting communities that I have become part of during my PhD, including but not limited to: parkrun, GoodGym, LGBT+ Fitness Class at Hamilton House, Project Awesome and Spin City.

My life has also changed in other ways. I would like to thank the wonderful folk at the University of Bristol LGBT+ Society, the Bristol SU Trans Students' Network, transcaf and Off the Record Bristol, for the wonderful support they have given me. And a particular thank you to Iris Dinu, for helping me find my direction in the first place.

Thank you to the many friends I have made or kept over the last 27 years, and the warmest thank you of all to my family, for the unconditional love, inspiration and support I have received my whole life. This wouldn't have been possible without you all.

\vspace{1cm}

\emph{This thesis is dedicated to my grandmothers, Stella Mae Lewis on my mother's side and Anne Moylett on my father's side, both of whom were immensely proud about their grandchild going to do a PhD. Wherever you two are now, I hope you are just as pleased today.}
